---
title: |  
      | Simulation and Optimization
      | ITAO 40150
header-includes:
   - \usepackage[dvipsnames]{xcolor}
output: pdf_document
fontsize: 12pt
urlcolor: CornflowerBlue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, comment = "")
```

# Office: 337 Mendoza

# Email: seth.berry@nd.edu

# Office Hours

Monday & Wednesday: 10:00 to 12:00

Tuesday & Thursday: 12:00 to 1:30

Most Fridays: 10:00 to 12:00

# Class Days and Time

Monday and Wednesday: B051

- Section 1: 2:00 - 3:15

- Section 2: 3:30 - 4:45

# Course Objectives

By the end of the course, you will be able to:

- Identify problems that can be solved using advanced quantitative techniques, like linear programming and simulation.

- Translate word-based optimization problems into mathematical expressions for further programming.

- Know when to apply specific types of optimization problems.

- Recognize how statistical distributions can be related to real-life situations.

- Program simulations using programming languages of your choice.

# Course Activities

## Readings

There is no *official* textbook for this course. There will, however, be readings posted for each topic. You are not committed to the readings, but they will certainly provide an expanded understanding of the content. 

## Class Exercises

After learning a topic, you will work individually in class to solve one or two questions related to the new topic. Although these exercises will not be graded, you will get much more out of them then listening to me talk about how to do these things.

## Homework

There will be three assignments which will be graded. These assignments and suggested due dates will be announced during the class. 

Please feel free to work together on the bronze level of homework, but each assignment needs to be your own work.  Putting your heads together to formulate an analytic attack plan is perfect (we all stands on the shoulders of giants), but copying and pasting text from each other is unacceptable. In other words, your code and words should not look like one of your classmate's. 

## Midterm

The midterm will require programming. However, the scope of the midterm is limited to only a few topics.

## Presentations

The final for this class will be a group presentation (and accompanying write-up). You will be graded on the in-class presentation, the write-up, and peer evaluations.

# Grading

Daily Work -- 50 points (16%)

Homework -- 90 points (29%)

Midterm Exam -- 70 points (23%)

Final Presentation -- 100 points (32%)

Total -- 310 points


\newpage

# Schedule

```{r, results='asis'}
library(dplyr)
data.frame(Week = c(1, "", 2, "", 3, "",
                    4, "", 5, "", 6, "", 
                    7, "", 8, ""),
          Date = c("10/12 (T)", 
                   "10/14 (Tr)",
                   "10/19 (T)", 
                   "10/21 (Tr)", 
                   "10/26 (T)", 
                   "10/28 (Tr)", 
                   "11/02 (T)", 
                   "11/04 (Tr)", 
                   "11/16 (T)", 
                   "11/18 (Tr)", 
                   "11/23 (T)",
                   "11/25 (Tr)",
                   "11/30 (T)",
                   "12/02 (Tr)", 
                   "12/07 (T)", 
                   "Final"), 
          Topic = c("Software Prep", 
                    "R Markdown",
                    "Fall", 
                    "Break",
                    "Linear Programming 1", 
                    "Linear Programming 2",
                    "Process Simulation 1", 
                    "Process Simulation 2", 
                    "Midterm Review",
                    "Midterm", 
                    "Integer Programming",
                    "Thanksgiving",
                    "Monte Carlo Simulation 2", 
                    "Project Day", 
                    "Project Day", 
                    "Presentations"), 
           Assignments = c("", "", 
                        "", "",
                        "HW1 Assigned", "",
                        "HW2 Assigned", "",
                        "", "",
                        "", "",
                        "HW3 Assigned",  "", 
                        "", "")) #%>% 
  stargazer::stargazer(., summary = FALSE, header = FALSE, title = "Tentative Schedule")
```