---
title: "Ali's Guide To Data Manipulation"
output:
  distill::distill_article:
    toc: true
    toc_float: true
    highlight: tango
    highlight_downlit: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = "")
```

Optimization seemed a little easy on the data manipulation side, but one particular problem had me a little shook. I was given a problem in a csv file and I needed to break it down into smaller pieces. Seemed easy to do, but I wanted to try it a few different ways: base R and dplyr.

## Data Read

```{r}
advert <- read.csv("https://raw.githubusercontent.com/saberry/SimOps/main/advertising_data.csv")

advert
```

It seems like that X column must have been row names within the sheet -- that's okay, I think I'll just give it a name:

```{r}
names(advert)[which(names(advert) == "X")] <- "rules"
```

I took the names of advert and then tested which of those names was strictly equal to "X" -- that is the name that I changed to rules.

Having that as a column should be pretty easy now.

## Filtering

In all my looking around on StackOverflow, it seems like there is always a base R way of doing things and a dplyr way of doing things. If I need to filter my data down, I'd like to see what that looks like both ways. I'm going to start with pulling the `c vector` out. When I look at the data, I know that the cvec is in the very last row. I know that I am going to need to drop the rules and demographic_total columns

### base R

```{r}
base_objective_function <- advert[advert$rules == "cost", ]
```

The row side of this is simple: give me the rows were advert$rules is strictly equal to "cost". 

### dplyr

Let's see how this same function would work in dplyr:

```{r}
library(dplyr)

dplyr_objective_function <- advert %>%
  filter(rules == "cost")
```

So `filter` seems pretty simple here. It isn't much different than what I do with base R, but it does have a nice verb to remind me about what I'm doing.

## Selecting Rows

### base R

```{r}
base_objective_function[, !(names(advert) %in% c("rules", "demographic_total"))]
```

Since I already filtered the rows, I can move to the column selection part; however, it seems a little trickier. I will ignore the `!` for now, so I can worry about what is in the parentheses. First I am running the names function on advert -- this just provides a vector of the names within the data frame. Then, I am using the `%in%` operator to match the names to the vector containing just "rules" and "demographic_total". This finds the names in advert that offer an exact match. This is what I would get:

```{r}
names(advert) %in% c("rules", "demographic_total")
```

A whole bunch of TRUEs and FALSEs.

When I throw the `!` in at the front, I am saying "give me the things that do not match" -- in other words, just return the columns that are equal to FALSE. This is a trick that I'll probably find very helpful in many data wrangling quests.

### dplyr

```{r}
dplyr_objective_function %>%
  select(cosmopolitan:country_living)
```

Since I created our `dplyr_objective_function` object already, I'll pipe to `select` -- there are a lot of different ways to select data here. In this select, I have included the range of variables from `cosmopolitan` to `country_living`; the `:` is working just like it typically does in R by providing a range.

## Filter And Select

I'd figure that there is a way to put all of these things together into "single" lines:

### base R

```{r}
base_objective_function <- advert[advert$rules == "cost", 
                                  !(names(advert) %in% 
                                      c("rules", "demographic_total"))] |>
  as.matrix()
```


### dplyr

```{r}
dplyr_objective_function <- advert %>%
  filter(rules == "cost") %>%
  select(cosmopolitan:country_living) %>% 
  as.matrix()
```

And I can! The dplyr version is definitely more "readable", but I feel like the base way requires less work. Purely curious, but I wonder if there is a speed difference:

```{r}
library(microbenchmark)

base_function <- function(){
  advert[advert$rules == "cost", 
         !(names(advert) %in% 
             c("rules", "demographic_total"))] |>
    as.matrix()
}

dplyr_function <- function(){
  advert %>%
    filter(rules == "cost") %>%
    select(cosmopolitan:country_living) %>% 
    as.matrix()
}

microbenchmark(base_function(), dplyr_function(), times = 500)
```

It seems like base is much faster...interesting.

<aside>
Speed differences are weird. For some problems, dplyr might be faster.  
Might also mean your base isn't as good as it could be.  
The data.table package reigns supreme when speed is at play.
</aside>

## Single Chunks For Solving

### base R

```{r}
library(linprog)

base_objective_function <- advert[advert$rules == "cost", 
                                  !(names(advert) %in% 
                                      c("rules", "demographic_total"))] |>
  as.matrix()
```

Since I only want the demographic totals, I can just use the dollar sign and then say give me the rows where rules does not equal "cost":

```{r}
base_constraint_margins <- advert$demographic_total[advert$rules != "cost"] |>
  as.matrix()

base_constraint_directions <- rep(">=", length(base_constraint_margins))

base_constraint_matrix <- advert[advert$rules != "cost", 
                                 !(names(advert) %in% 
                                     c("rules", "demographic_total"))] |>
  as.matrix()

base_res <- solveLP(base_objective_function, 
                    base_constraint_margins, 
                    base_constraint_matrix, 
                    maximum = FALSE, 
                    const.dir = base_constraint_directions)

base_solution <- base_res$solution
```


### dplyr

```{r}
dplyr_objective_function <- advert %>%
  filter(rules == "cost") %>%
  select(cosmopolitan:country_living) %>% 
  as.matrix()
```

It turns out that dplyr is going to try to keep all the names within my margins; this sounds great, but linprog will cry if those dimnames are there. To that end, I will have to `unname` at the end of my next pipe.

```{r}
dplyr_constraint_margins <- filter(advert, rules != "cost") %>%
  select(demographic_total) %>% 
  as.matrix() %>%
  unname()

# No special rep -- the base R rep is all anyone needs:
dplyr_constraint_directions <- rep(">=", length(dplyr_constraint_margins))

dplyr_constraint_matrix <-   advert %>% 
  filter(rules != "cost") %>%
  select(cosmopolitan:country_living) %>%
  as.matrix()

dplyr_res <- solveLP(dplyr_objective_function, 
                     dplyr_constraint_margins, 
                     dplyr_constraint_matrix, 
                     maximum = FALSE, 
                     const.dir = dplyr_constraint_directions)

dplyr_solution <- dplyr_res$solution
```


## The Final Straw: Bind/Merge/Join

I've got solutions either way, but what I really want is to be able to put those what I achieved back into my original data.

I'll just use the base results, because they are all the same:

```{r}
str(base_res)
```

I see that base_res is a list, that contains a mix of different things. It looks like the object I want is called `con`:

```{r}
base_res$con
```

That gives me the data frame, so I can probably just toss another `$` in there:

```{r}
achieved_values <- base_res$con$actual
```

No doubt that there are many ways to handle this task, so I'll try some out:

```{r, eval = FALSE}
advert$base_achieved_values <- achieved_values
```

> Error in `$<-.data.frame`(`*tmp*`, base_achieved_values, value = c(210,  :   
> replacement has 6 rows, data has 7

This won't work -- I don't have a match in the length of `achieved_values` and the nrow() of advert.

Let me see if dplyr works:

```{r, eval = FALSE}
advert %>%
  mutate(dplyr_achieved_values = achieved_values)
```

> Error: Problem with `mutate()` column `dplyr_achieved_values`.  
> i `dplyr_achieved_values = achieved_values`.  
> i `dplyr_achieved_values` must be size 7 or 1, not 6.  
> Run `rlang::last_error()` to see where the error occurred.  

Same problem, just a slightly different message.

I wonder if I can bind them:

```{r, eval = FALSE}
cbind(advert, achieved_values)
```

> Error in data.frame(..., check.names = FALSE) : 
> arguments imply differing number of rows: 7, 6

Maybe dplyr will work for this one:

```{r, eval = FALSE}
bind_cols(advert, achieved_values)
```

> Error: Can't recycle `..1` (size 7) to match `..2` (size 6).  

I have been forsaken. Every method that I would usually use to add a column to a data frame will not work. 

After some time poking around, I figured out a way to handle the situation: merge/join!

I can start with taking those results and putting them into a data frame:

```{r}
achieved_results_df <- data.frame(achieved_values = achieved_values)
```

Next, I can take the rules from our original data and put them into our new data frame:

```{r}
advert_rules <- advert$rules[advert$rules != "cost"]

achieved_results_df$rules <- advert_rules
```

Awesome, I just created an object called `advert_rules` that was just the rules column with the cost row dropped out.

I could have also used the mutate function:

```{r}
achieved_results_df %>%
  mutate(dplyr_rules = advert_rules)
```

Certainly no difference in the results!

Finally, we can try two different methods for bringing these two data frames together:

### base R

In base R, we have a function called `merge`:

```{r}
merge(x = advert, y = achieved_results_df, # These are the data frames I am merging
      by.x = "rules", by.y = "rules", # These are the names from x&y that I am using
      all.x = TRUE, all.y = FALSE) # I want to keep everything in advert
```

You likely noticed that we set both `by.x` and `by.y` -- this is most useful when the column names you want to join are different. Since our columns have the same names, we could also have done this:

```{r}
merge(x = advert, y = achieved_results_df, 
      by = "rules",
      all.x = TRUE, all.y = FALSE) 
```


### dplyr

If we live in the SQL world, we just did a `left_join`, which is exactly what we can do with dplyr:

```{r}
left_join(x = advert, y = achieved_results_df, 
          by = c("rules" = "rules"))
```

Same `by` issue applies here:

```{r}
left_join(x = advert, y = achieved_results_df, 
          by = "rules")
```

