---
title: "LP Exercies"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


## Working Together

```{r}
library(linprog)

# First, we need to read in data from the web. The row.names = 1 argument
# is saying to set the first column in the data as the row names.

advert <- read.csv("https://raw.githubusercontent.com/saberry/SimOps/main/advertising_data.csv", 
                   row.names = 1)

# Everything we need is in that advert data frame, but we have to break
# it into smaller chunks: the cvec, bvec, and amat. The easiest
# way to make that happen is to find the objects that you want to
# remove from the entire data frame.

# Easiest to break these functions down from the inside. 
# The names function will return all of the column names
# within the advert data -- it just prints out a vector
# of all the variable names. When that is wrapped in the
# which function, it answers the question of "Which
# of those variables names is an exact match to the word
# "demographic_total; the exact matching column number if
# 10.

dropDemos <- which(names(advert) == "demographic_total")

# Now that we now the demographic column is identified, we
# can drop that column using -c(dropDemos) -- literally saying
# to drop the 10th column. For the cvec, we are using the
# cost row. Putting it together, we will select the "cost" row, 
# drop the demographic column and convert the result to a matrix.

objectiveFunction <- as.matrix(advert["cost", -c(dropDemos)])

# Now, we need to drop out the cost row. We can engage in a very
# similar process to what we did earlier, but instead we can use
# the rownames function. 

dropCost <- which(rownames(advert) == "cost")

# The margins of the problem rest in the demographic column (i.e., 
# we need to have at least this many exposures across the 6
# demographic groups). We will drop the cost row -- -c(dropCost) --
# and then select the demographic_total column.

constraintMargins <- as.matrix(advert[-c(dropCost), "demographic_total"])

# Finally, we can pull out the A matrix by dropping the cost row and
# dropping the demographic column.

constraintMatrix <- as.matrix(advert[-c(dropCost), -c(dropDemos)])

res <- solveLP(objectiveFunction, constraintMargins, constraintMatrix, 
              maximum = FALSE, const.dir = rep(">=", length(constraintMargins)))

res$opt

res$solution

res$con
```

### Keeping Track Of Variables

### Lower and Upper Bounds


## Solve this problem

$$
\begin{aligned}
\text{Maximize:}\\
\text{revenue} = 40_{oil} + 7.5_{flower} \\
\text{Subject to:} \\
\text{needs} = 30_{oil} + 3_{flower} \leq 8960 \\
\text{oil}, \text{flower} \geq 0
\end{aligned}
$$

```{r}
library(linprog)

# The bottom line, "oil,flower >= 0" is implicit in the problem and never
# needs specified.

objectiveFunction <- c(40, 7.5) # This is the top line (cvec) of the equation

constraintValues <- c(8960) # We only have one real margin (bvec) to handle here.

constraintDirections <- c("<=")

constraintMatrix <- rbind(c(30, 3)) # Only one row in the constraint matrix (A mat)

linprog::solveLP(cvec = objectiveFunction, 
                 bvec = constraintValues, 
                 Amat = constraintMatrix, 
                 maximum = TRUE, 
                 const.dir = constraintDirections)
```

In total, we could make 22400 dollars, by making 0 units of oil and 2986.67 units of flower.

### Adding Upper and Lower Bounds

Produce at least 50 grams of oil and make note of what happens.

This just becomes an extra constraint to the model (we are saying that we need to make at least 50 units of oil, but nothing specified for flower):

$$
\begin{aligned}
\text{Maximize:}\\
\text{revenue} = 40_{oil} + 7.5_{flower} \\
\text{Subject to:} \\
\text{needs} = 30_{oil} + 3_{flower} \leq 8960 \\
\text{oil_needs} = 1_{oil} + 0_{flower} \geq 50 \\
\end{aligned}
$$

We are going to do this two ways: both will yield the same result, so use whichever makes the most sense to you.

#### linprog

```{r}
objectiveFunction <- c(40, 7.5) # Unless we add another variable, nothing changes here.

constraintValues <- c(8960, 50) # Just adding the extra information into out bvec

constraintDirections <- c("<=", ">=") # One more direction to deal with.

# We have an extra row, with the 1,0 combination for oil.
constraintMatrix <- rbind(c(30, 3), 
                          c(1, 0)) 

linprog::solveLP(cvec = objectiveFunction, 
                 bvec = constraintValues, 
                 Amat = constraintMatrix, 
                 maximum = TRUE, 
                 const.dir = constraintDirections)
```

That works just fine and will always work. Let's see how that works with ROI.

Remember, ROI has functions in place to make this an easier process:


```{r}
library(ROI)

# You will notice that all of these are the original objects.

objectiveFunction <- c(40, 7.5) # This is the top line (cvec) of the equation

constraintValues <- c(8960) # We only have one real margin (bvec) to handle here.

constraintDirections <- c("<=")

constraintMatrix <- rbind(c(30, 3)) # Only one row in the constraint matrix (A mat)

# ROI makes us create the constraint object using this L_constraint function.
# The argument are just as they appear in the mathematic specification:

linear_constraints <- L_constraint(L = constraintMatrix, 
                                   dir = constraintDirections, 
                                   rhs = constraintValues)

# Now, we can specify our lower bound for the oil variable:

# In our model, we have two variables: oil and flower. 
# If we put a number order to those variables we have:
# oil = 1
# flower = 2
# We only want to set a lower bound for oil, so we can
# specify an li (i.e., lower index value) to 1 and
# set the lb (i.e., the lower bound value) to 50.

oil_lower_bound <- V_bound(li = c(1), lb = c(50), nobj = 2)

# This only creates the model:
model_create <- OP(objective = objectiveFunction, 
                   constraints = linear_constraints, 
                   bounds = oil_lower_bound, 
                   maximum = TRUE)

# This solves the model:
model_solved <- ROI_solve(model_create)

# This gives us the values for oil and flower (i.e., x = 50 & y = 2486.667)
solution(model_solved)

#And now we have the total amount made:
solution(model_solved, "objval")
```

The return is exactly the same that we got from linprog. Again, I leave it to you to determine which you find easier.


Keep your lower bound for oil, and also specify an upper bound to keep flower less than or equal to 2000. How does your solution change?

Let's update the notation:

$$
\begin{aligned}
\text{Maximize:}\\
\text{revenue} = 40_{oil} + 7.5_{flower} \\
\text{Subject to:} \\
\text{needs} = 30_{oil} + 3_{flower} \leq 8960 \\
\text{oil_needs} = 1_{oil} + 0_{flower} \geq 50 \\
\text{flower_cap} = 0_{oil} + 1_{flower} \leq 2000 \\
\end{aligned}
$$

```{r}
objectiveFunction <- c(40, 7.5) # Still nothing changes here.

# Now, we have a third constraint: the upper bound for flower:
constraintValues <- c(8960, 50, 2000) 

# And that upper bound has a <= constraint:
constraintDirections <- c("<=", ">=", "<=") 

# We have an extra row, with the 1,0 combination for oil.
# And also the 0, 1 combination for flower:
constraintMatrix <- rbind(c(30, 3), 
                          c(1, 0), 
                          c(0, 1)) 

linprog::solveLP(cvec = objectiveFunction, 
                 bvec = constraintValues, 
                 Amat = constraintMatrix, 
                 maximum = TRUE, 
                 const.dir = constraintDirections)
```

And our flower has been capped at 2000, as we have specified.

Now, we can try it with ROI:

```{r}
objectiveFunction <- c(40, 7.5) 

constraintValues <- c(8960) 

constraintDirections <- c("<=")

constraintMatrix <- rbind(c(30, 3)) 

linear_constraints <- L_constraint(L = constraintMatrix, 
                                   dir = constraintDirections, 
                                   rhs = constraintValues)

# We already have our lower bound for oil (oil = index value 1), 
# so we can add an upper bound for flower. Remember, flower is the
# second variable, so it has an index value of 2. 

oil_lower_bound <- V_bound(li = c(1), lb = c(50), 
                           ui = c(2), ub = c(2000),
                           nobj = 2)

# This only creates the model:
model_create <- OP(objective = objectiveFunction, 
                   constraints = linear_constraints, 
                   bounds = oil_lower_bound, 
                   maximum = TRUE)

# This solves the model:
model_solved <- ROI_solve(model_create)

# This gives us the values for oil and flower (i.e., x = 50 & y = 2486.667)
solution(model_solved)

#And now we have the total amount made:
solution(model_solved, "objval")
```

Again, we have the exact same output!