# Process Simulation

Ali was absolutely smoked (no pun intended) after handling all of that optimization -- hopefully some more standard modeling would come through. High hopes are always short lived, though, and Ali was thrown right back into the dark arts of Operations-based research.

Alex was the resident expert of simulations of all kind, so Ali knew an ally was there. 

> Hi Ali

## Discrete Event Simulation



## Queueing Theory

Queues have their own notation (Kendall's notation). 

$A/B/C/D$

Where:

$A = arrival\, process$

$B = service\, process$

$C = server\, number$

$D = que\, capacity$

$M/D/k$

$M/M/k$

*M* generally stands for Markov or Exponential

<aside>
You might also see *M* described as being from a Poisson process
</aside>

*D* is deterministic: all jobs require a fixed amount of time.

*k* is the number of servers/workers/etc.

Both of these are generally assumed to have an infinite buffer.

If a queue is $M/D/k$, we can easily compute some helpful statistics (we don't need any fancy software to help us).

$\lambda$ = arrival rate

$\mu$ = service rate

$\rho = \frac{\lambda}{\mu}$ = utilization

Average number of entities in the system is:

$$ L = \rho + \frac{1}{2}\Bigg(\frac{\rho^2}{1 - \rho}\Bigg) $$

Average number in queue:

$$ L_Q = \frac{1}{2}\Bigg(\frac{\rho^2}{1 - \rho}\Bigg) $$

Average system waiting time:

$$ \omega = \frac{1}{\mu}+\frac{\rho}{2\mu(1 - \rho)} $$

Average waiting time in queue:

$$ \omega_Q = \frac{\rho}{2\mu(1 - \rho)} $$

## Process Flow Map

How does the flow for a bank typically look:

1.  A customer enters the bank through the door.

2.  The customer will either go directly to a teller or wait in line for the teller.

3.  Teller will serve the customer.

4.  The customer has been served and will exit the bank.

```{r}
library(DiagrammeR)

grViz("
digraph {
  graph [overlap = true, fontsize = 10, rankdir = LR]
  
  node [shape = box, style = filled, color = black, fillcolor = aliceblue]
  A [label = 'Door']
  B [label = 'Line']
  C [label = 'Teller']
  D [label = 'Served Customer']

  A->B B->C C->D
}
")
```

## Distributions

### Normal Distribution

For our normal distribution, we know the $\mu$ and $\sigma$.

```{r, echo = FALSE}
library(dplyr)

library(ggplot2)

data.frame(x = rnorm(10000, mean = 2.5, sd = .5)) %>% 
  ggplot(., aes(x = x, y = ..density..)) +
  geom_histogram(color = "black", fill = "white", bins = 15) +
  geom_density(color = "#ff5500", size = 1.25) +
  theme_minimal() +
  labs(x = "Call Center Wait Times", y = "Frequency")
```


### Exponential Distribution

We can only know one thing about the exponential distribution: $\mu$ (also expressed as a rate).

```{r, echo = FALSE}
data.frame(x = rexp(10000, rate = 2.5)) %>% 
  ggplot(., aes(x = x)) +
  geom_histogram(color = "black", fill = "white", bins = 15) +
  theme_minimal() +
  labs(x = "Call Center Wait Times", y = "Frequency")
```

### Poisson Distribution

The poisson is an interesting distribution -- it tends to deal with count-related variables. It tells us the probability of a count occuring.  We know its $\lambda$ (again, a rate).

<aside>

$\lambda$ is just a fancy way of saying the average number of events, or the incidence rate.

</aside>

```{r, echo = FALSE}
data.frame(x = rpois(10000, lambda = 4)) %>% 
  ggplot(., aes(x = x)) +
  geom_bar(color = "black", fill = "white") +
  theme_minimal() +
  labs(x = "Customers In Line", y = "Frequency")
```

### Uniform Distribution

While people tend to think about the Gaussian distribution as the most vanilla of all distributions, it really is not -- I would say that distinction belongs to the uniform distribution. We don't even get any fancy Greek letters, just a minimum and a maximum. Why? Because knowing the min and max will tell us that there is an equal probability of drawing a value anywhere within that range.

```{r, echo = FALSE}
data.frame(x = runif(10000, min = 1, max = 10)) %>% 
  round(.) %>% 
  ggplot(., aes(x = x)) +
  geom_bar(color = "black", fill = "white") +
  theme_minimal() +
  labs(x = "Customers In Line", y = "Frequency")
```

## Performance

The *service level* for each simulation is the fraction of the demand that is satisfied.

$$ Entrance  \: Service \: Level = \frac{Objects \: Entering}{Objects \: Entering + Objects \: Unable \: To \: Enter}$$

Here, we are looking at the number of people who wanted to join the process, but could not. If we have a service level of 1, then 100% of objects were able to get into the process. A service level of .5 would indicate that only 50% of objects were able to enter.

The *overall mean service level* of the process is the mean of the service levels calculated from each simulation.

The *mean cycle time* at a buffer is the mean amount of time an object takes to move through the buffer during a simulation.

The *overall mean cycle time* at a buffer is the mean of the mean cycle time of the buffer for each simulation.

<aside>
You will see different words for lines: buffers and queues. Just know that they are used interchangeably. 
</aside>

## The Dispensary

The interarrival times for customers follows an exponential distribution with $\mu = 2 \, minutes$.

The dispensary cannot hold any more than 8 people, for safety reasons. If a person arrives when the line is full, that person will not get in line.

The bud tender's service time can be approximated by a normal distribution with $\mu = 2.4 \, minutes$ and $\sigma = .5 \, minutes$.

```{r}
library(DiagrammeR)
grViz("
digraph {
  graph [overlap = true, fontsize = 10, rankdir = LR]
  
  node [shape = box, style = filled, color = black, fillcolor = aliceblue]
  
  A [label = 'Door']
  B [label = 'Line']
  C [label = 'Teller 1']
  D [label = 'Teller 2']
  E [label = 'Served']
  
  A->B 
  B->C
  B->D
  C->E
  D->E
}
")

```