# Process Simulation

Ali was absolutely smoked (no pun intended) after handling all of that optimization -- hopefully some more standard modeling would come through. High hopes are always short lived, though, and Ali was thrown right back into the dark arts of Operations-based research.

During an "all-hands" meeting, Ali had the chance to listen to Rene, the Director of Retail Analytics, talk about store efficiency. Rene explained the process that typically happens. 

"Our average store front is pretty small and we need to be careful about how many people are inside at any one time", Rene started. "We can't have more than 8 people waiting in the lobby before their ID's are checked", and Rene continued, "we don't want to turn people away, so we need to get more efficient in ID checks and bud-tending". 

Ali was feeling good after some success and wasn't afraid to ask some questions -- "Can you explain the process to me?", Ali asked. 

"Sure", Rene said. "It is pretty easy, people walk in the door and wait for their ID to be checked." 

"Once their ID has been checked they can meet with one of our bud-tenders -- they pick their poison and then pay."

Ali was drawing the process out and made sure that it was correct:

"Seem about right?", Ali asked.

```{r}
library(DiagrammeR)

grViz("
digraph {
  graph [overlap = true, fontsize = 10, rankdir = LR]
  
  node [shape = box, style = filled, color = black, fillcolor = aliceblue]
  A [label = 'ID Check Line']
  B [label = 'ID Check']
  C [label = 'Bud Tender Line']
  D [label = 'Bud Tender']
  E [label = 'Pay']

  A->B B->C C->D D->E
}
")
```

"That's right", Rene said.

"How many people are checking IDs and how many bud tenders do you have?", Ali asked. 

"It kinda depends", Rene said, "but it is usually just one person checking IDs and usually two bud tenders."

Ali had one more question: "How long do each of those steps take?"

"Uhhhhh...I'll have to ask around and get back with you", Rene noted.

"Most excellent", Ali thought, "that will give me some time to chat with Alex." 

## Discrete Event Simulation

When Ali finally caught up with Alex, Alex was only too happy to share some of the finer points on process simulation. First, Alex made note that the particular type of simulation under conversation wasn't just a process simulation, but was something called *discrete event simulation* (DES) -- events are individual processes and some items goes through a series of those individual processes. 

"It has roots in manufacturing, but some many things in life can be modeled through DES", Alex said. 

"Let's start with something horribly boring...lines."

## Queueing Theory

"There is a whole field of study regarding lines", Alex said, "and it is called *queueing theory*."

"We don't have to get crazy, but there is this thing called Kendall's Notation...it just describes the particular parts of how lines form and the distributions that guide them."

Ali thought, "The theory of lines...how do some people ever find love."

Alex could almost feel Ali's thought and said, "It really is more interesting than it sounds."

"Check this out!", and Alex was off to the races.

$A/B/C/D$

Where:

$A = arrival\, process$

$B = service\, process$

$C = server\, number$

$D = que\, capacity$

$M/D/k$

$M/M/k$

*M* generally stands for Markov or Exponential

*D* is deterministic: all jobs require a fixed amount of time.

*k* is the number of servers/workers/etc.

"Both of these are generally assumed to have an **infinite queue**...that is important to remember."

"If a queue is $M/D/k$, we can easily compute some helpful statistics."

$\lambda$ = arrival rate

$\mu$ = service rate

$\rho = \frac{\lambda}{\mu}$ = utilization

Average number of entities in the system is:

$$ L = \rho + \frac{1}{2}\Bigg(\frac{\rho^2}{1 - \rho}\Bigg) $$

Average number in queue:

$$ L_Q = \frac{1}{2}\Bigg(\frac{\rho^2}{1 - \rho}\Bigg) $$

Average system waiting time:

$$ \omega = \frac{1}{\mu}+\frac{\rho}{2\mu(1 - \rho)} $$

Average waiting time in queue:

$$ \omega_Q = \frac{\rho}{2\mu(1 - \rho)} $$

"The equations are not the important part here", Alex said, "but the idea that the equation captures is critical."

## Performance

The *service level* for each simulation is the fraction of the demand that is satisfied.

$$ Entrance  \: Service \: Level = \frac{Objects \: Entering}{Objects \: Entering + Objects \: Unable \: To \: Enter}$$

Here, we are looking at the number of people who wanted to join the process, but could not. If we have a service level of 1, then 100% of objects were able to get into the process. A service level of .5 would indicate that only 50% of objects were able to enter.

The *overall mean service level* of the process is the mean of the service levels calculated from each simulation.

The *mean cycle time* at a buffer is the mean amount of time an object takes to move through the buffer during a simulation.

The *overall mean cycle time* at a buffer is the mean of the mean cycle time of the buffer for each simulation.

<aside>
You will see different words for lines: buffers and queues. Just know that they are used interchangeably. 
</aside>

## The Dispensary

The interarrival times for customers follows an exponential distribution with a rate of 1 person every 1.5 minutes.

The dispensary cannot hold any more than 8 people, for safety reasons. If a person arrives when the line is full, that person will not get in line.

The ID check is approximately normal with $\mu=15 \, seconds$ and $\sigma = 3 \, seconds$. Once a person has their ID checked, they can sit in the lobby and there are 10 seats in the lobby.

The bud tender's service time is $\mu = 2.4 \, minutes$ and $\sigma = .5 \, minutes$.

Paying generally follows a uniform distribution, with a minimum of 5 seconds and a maximum of 15 seconds.

```{r}
grViz("
digraph {
  graph [overlap = true, fontsize = 10, rankdir = LR]
  
  node [shape = box, style = filled, color = black, fillcolor = aliceblue]
  A [label = 'ID Check Line']
  B [label = 'ID Check']
  C [label = 'Bud Tender Line']
  D [label = 'Bud Tender']
  E [label = 'Pay']

  A->B B->C C->D D->E
}
")
```

### ClassOverflow

We will need the `simmer` package for our simulation:

```{r, eval = FALSE}
install.packages("simmer")
```

Once we have `simmer` installed, we need to load it:

```{r, echo = TRUE}
library(simmer)
```

Let's start by defining a customer's trajectory. First, we will provide a name for `trajectory()`. 

```{r, echo = TRUE, eval = FALSE}
customer <- trajectory("Customer path")
```

Next, we need to initiate a start time with `set_attribute()` -- we will use `now()` to specify our not-yet-created dispensary object.

```{r, echo = TRUE, eval = FALSE}
customer <- trajectory("Customer path") |>
  set_attribute("start_time", function() {now(dispensary)})
```

After establishing our time, the next step for a customer is to `seize()` the "teller" (which we will define later).

```{r, echo = TRUE, eval = FALSE}
customer <- trajectory("Customer path") |>
  set_attribute("start_time", function() {now(dispensary)}) |>
  seize("id_check")
```

Now things start to get tricky. We need to use `timeout()` to specify how long a customer is using the id check -- this is the check's average working time. 

We can specify how long an id check is seized (i.e., how long the check is working)  -- we provide a distribution with the appropriate values.

```{r, echo = TRUE, eval = FALSE}
customer <- trajectory("Customer path") %>% 
  set_attribute("start_time", function() {now(dispensary)}) |>
  seize("id_check") |>
  timeout(function() {rnorm(n = 1, mean = 15/60, sd = 3/60)})
```

After a customer spends time with the teller, the customer releases the counter.

```{r, echo = TRUE, eval = FALSE}
customer <- trajectory("Customer path") %>% 
  set_attribute("start_time", function() {now(dispensary)}) |>
  seize("id_check") |>
  timeout(function() {rnorm(n = 1, mean = 15/60, sd = 3/60)}) |>
  release("id_check")
```

From there, we can add the additional resources to our model:

```{r, echo = TRUE, eval = FALSE}
customer <- trajectory("Customer path") %>% 
  set_attribute("start_time", function() {now(dispensary)}) |>
  seize("id_check") |>
  timeout(function() {rnorm(n = 1, mean = 15/60, sd = 3/60)}) |>
  release("id_check") |>
  seize("bud_tender") |>
  timeout(function() {rnorm(n = 1, mean = 2.4, sd = .5)}) |>
  release("bud_tender") |>
  seize("payment") |>
  timeout(function() {runif(n = 1, min = 5/60, max = 15/60)}) |>
  release("payment")
```


This is all we need to do for a customer, so now we can turn our attention to the dispensary.

Our dispensary is going to provide the environment that houses our trajectory. So, we can start by creating an environment with `simmer()`:

```{r, echo = TRUE, eval = FALSE}
dispensary <- simmer("dispensary")
```

Once we have our simulation environment defined, we can add resources to it with the aptly-named `add_resources()` function. This is where we will specify what is being seized by our customer. We need to provide some additional information to our resource: `capacity` and `queue_size`. 

```{r, echo = TRUE, eval = FALSE}
dispensary <- simmer("dispensary") |>
  add_resource("id_check", capacity = 1, queue_size = 8) |>
  add_resource("bud_tender", capacity = 2, queue_size = 10) |>
  add_resource("payment", capacity = 2) 
```

To this point, we have our customer behavior (how they move through our process) and information about our work stations. The last detail is the inter-arrival time, which we can specify with `add_generator()`. It works in very much the same way that `timeout()`, in that we are specifying a distribution. The `rexp` function in R takes a rate. If we remember that, on average, one person comes into the dispensary every two minutes, we can define our rate as $\frac{1}{2}$.

Try this: mean(rexp(n = 10000, rate = 1/2))

```{r, echo = TRUE, eval = FALSE}
dispensary <- simmer("dispensary") |>
  add_resource("id_check", capacity = 1, queue_size = 8) |>
  add_resource("bud_tender", capacity = 2, queue_size = 10) |>
  add_resource("payment", capacity = 2) |>
  add_generator("Customer", customer, function() {
    c(0, rexp(n = 100, rate = 1/1.5), -1)
  })
```

Now we can simmer::run our simulation; we just need to provide a time value for the `until` argument. Let's say we want to run this simulation for 2 hours.

```{r, echo = TRUE, eval = FALSE}
simmer::run(dispensary, until = 120)
```

If we put it together, here is what we have:

```{r, echo = TRUE}
customer <- trajectory("Customer path") %>% 
  set_attribute("start_time", function() {now(dispensary)}) |>
  seize("id_check") |>
  timeout(function() {rnorm(n = 1, mean = 15/60, sd = 3/60)}) |>
  release("id_check") |>
  seize("bud_tender") |>
  timeout(function() {rnorm(n = 1, mean = 2.4, sd = .5)}) |>
  release("bud_tender") |>
  seize("payment") |>
  timeout(function() {runif(n = 1, min = 5/60, max = 15/60)}) |>
  release("payment")

dispensary <- simmer("dispensary") |>
  add_resource("id_check", capacity = 1, queue_size = 8) |>
  add_resource("bud_tender", capacity = 2, queue_size = 10) |>
  add_resource("payment", capacity = 2) |>
  add_generator("Customer", customer, function() {
    c(0, rexp(n = 100, rate = 1/1.5), -1)
  })

simmer::run(dispensary, until = 120)
```

Finally, we can start to look at our data:

```{r, echo = TRUE}
result <- get_mon_arrivals(dispensary)

head(result)
```

Let's calculate a few things. First, let's how many people made it through: 

```{r, echo = TRUE}
nrow(result[result$finished == TRUE, ])
```

Now we can check our service level:

```{r, echo = TRUE}
nrow(result[result$finished == TRUE, ]) / nrow(result)
```

The `nrow` function will tell us how many rows are in the data. In the numerator, we filtered those rows were finished was equal to TRUE (giving us the number of people who made it into the system).

Now we need to calculate how long each person was in line. 

```{r, echo = TRUE}
result$wait_time <- result$end_time - result$start_time - result$activity_time
```

Now, we can find the average wait time. We only want to do it for those who actually made it into the system though!

```{r, echo = TRUE}
completeOnly <- result[result$finished == TRUE, ]

mean(completeOnly$wait_time)
```

That gives us all of the information that we need for this bank configuration.

But...that is just one simulation. We really need to run this many times to get an idea about the distribution of outcomes.

We have a few choices. One choice is that we just replicate our procedure a certain number of times:

```{r, eval = FALSE, echo = TRUE}
sim50Runs <- replicate(50, expr = {
  customer <- trajectory("Customer path") %>% 
    set_attribute("start_time", function() {now(dispensary)}) |>
    seize("id_check") |>
    timeout(function() {rnorm(n = 1, mean = 15/60, sd = 3/60)}) |>
    release("id_check") |>
    seize("bud_tender") |>
    timeout(function() {rnorm(n = 1, mean = 2.4, sd = .5)}) |>
    release("bud_tender") |>
    seize("payment") |>
    timeout(function() {runif(n = 1, min = 5/60, max = 15/60)}) |>
    release("payment")
  
  dispensary <- simmer("dispensary") |>
    add_resource("id_check", capacity = 1, queue_size = 8) |>
    add_resource("bud_tender", capacity = 2, queue_size = 10) |>
    add_resource("payment", capacity = 2) |>
    add_generator("Customer", customer, function() {
      c(0, rexp(n = 100, rate = 1/1.5), -1)
    })
  
  simmer::run(dispensary, until = 120)
  
  result <- get_mon_arrivals(bank)
  
}, simplify = FALSE)
```

We can extend this idea into something a bit more complex:

```{r}
purrr::map_df(1:100, ~{
  customer <- trajectory("Customer path") %>% 
    set_attribute("start_time", function() {now(dispensary)}) |>
    seize("id_check") |>
    timeout(function() {rnorm(n = 1, mean = 15/60, sd = 3/60)}) |>
    release("id_check") |>
    seize("bud_tender") |>
    timeout(function() {rnorm(n = 1, mean = 2.4, sd = .5)}) |>
    release("bud_tender") |>
    seize("payment") |>
    timeout(function() {runif(n = 1, min = 5/60, max = 15/60)}) |>
    release("payment")
  
  dispensary <- simmer("dispensary") |>
    add_resource("id_check", capacity = 1, queue_size = 8) |>
    add_resource("bud_tender", capacity = 2, queue_size = 10) |>
    add_resource("payment", capacity = 2) |>
    add_generator("Customer", customer, function() {
      c(0, rexp(n = 100, rate = 1/1.5), -1)
    })
  
  simmer::run(dispensary, until = 120)
  
  result <- get_mon_arrivals(dispensary)
  
  result$run <- .x
  
  result
})
```

Next, we can see what things might look like if we change parts of the process:

```{r}
purrr::map2_df(.x = 1:100, .y = runif(100, min = 1, max = 2), ~{
    customer <- trajectory("Customer path") %>% 
    set_attribute("start_time", function() {now(dispensary)}) |>
    seize("id_check") |>
    timeout(function() {rnorm(n = 1, mean = 15/60, sd = 3/60)}) |>
    release("id_check") |>
    seize("bud_tender") |>
    timeout(function() {rnorm(n = 1, mean = 2.4, sd = .5)}) |>
    release("bud_tender") |>
    seize("payment") |>
    timeout(function() {runif(n = 1, min = 5/60, max = 15/60)}) |>
    release("payment")
  
  dispensary <- simmer("dispensary") |>
    add_resource("id_check", capacity = 1, queue_size = 8) |>
    add_resource("bud_tender", capacity = 2, queue_size = 10) |>
    add_resource("payment", capacity = 2) |>
    add_generator("Customer", customer, function() {
      c(0, rexp(n = 100, rate = 1/.y), -1)
    })
  
  simmer::run(dispensary, until = 120)
  
  result <- get_mon_arrivals(dispensary)
  
  result$run <- .x
  
  result$arrival <- .y
  
  result
})
```

And a function to show:

```{r}
employee_change_mod <- purrr::map2_df(.x = 1:100, .y = sample(1:3, 100, TRUE, c(.1, .7, .2)), ~{
  customer <- trajectory("Customer path") %>% 
    set_attribute("start_time", function() {now(dispensary)}) |>
    seize("id_check") |>
    timeout(function() {rnorm(n = 1, mean = 15/60, sd = 3/60)}) |>
    release("id_check") |>
    branch(function() {sample(1:2, 1, prob = c(.8, .2))}, 
           # The "continue" indicates if the branches should continue through
           # the trajectory once the branch trajectory is completed.
           continue = c(TRUE, TRUE), 
           trajectory() %>%
             seize("bud_tender") |>
             timeout(function() {rnorm(n = 1, mean = 2.4, sd = .5)}) |>
             release("bud_tender"), 
           trajectory() |> 
             seize("devices") |>  
             timeout(function() {rnorm(1, 5, 1)}) |>  
             release("devices")) |>
    seize("payment") |>
    timeout(function() {runif(n = 1, min = 5/60, max = 15/60)}) |>
    release("payment")
  
  dispensary <- simmer("dispensary") |>
    add_resource("id_check", capacity = 1, queue_size = 8) |>
    add_resource("bud_tender", capacity = .y, queue_size = 10) |>
    add_resource("devices", capacity = 1, queue_size = 10) |> 
    add_resource("payment", capacity = 2) |>
    add_generator("Customer", customer, function() {
      c(0, rexp(n = 100, rate = 1/.y), -1)
    })
  
  simmer::run(dispensary, until = 120)
  
  result <- get_mon_arrivals(dispensary)
  
  result$run <- .x
  
  result$employees <- .y
  
  result
})

sapply(split(employee_change_mod, employee_change_mod$employees), function(x) {
  finished <- x[x$finished == TRUE, ]
  nrow(finished) / nrow(x)
})

employee_change_mod$cycleTime <- employee_change_mod$end_time - employee_change_mod$start_time

aggregate(employee_change_mod$cycleTime, 
          by = list(employee_change_mod$employees), 
          mean)

library(ggplot2)

ggplot(employee_change_mod, aes(cycleTime)) +
  geom_histogram(bins = 10) +
  facet_wrap(vars(employees)) +
  theme_minimal()
```

## Manufacturing

In manufacturing, one of the more important metrics we deal with is *throughput* (the total number of units produced). 

Variability, while not something we can always measure/predict, can come from a few sources.

- Machine/worker processing times

- Output quality

- Demand

- Supply

Controlling variability can help to increase throughput.

### An Example

- In a factory, we have five work stations arranged in a line (WS1 through WS5).

- Each work station is a machine with one operator.

  - WS1 (bandsaw): $\mu = 10;\sigma = 1$ (normal)
  
  - WS2 (contouring): $\mu = 5;\sigma = 2$ (normal)
  
  - WS3 (rough sanding): $min = 5;max = 15$ (uniform)
  
  - WS4 (finish sanding): $min = 10;max = 15$ (uniform)
  
  - WS5 (finish): $\mu = 10;\sigma = 2.5$ (normal)

- The work stations process one product that must move sequentially.

- Each work station has its own processing time.

- Product cannot be *stacked*

  - If WS3 has finished a unit, it cannot be passed onto WS4 if WS4 is still working on a product.
  
- We are looking at an 8 hour shift.

### Improvements

What can we do to improve our throughput?

```{r, echo = TRUE}
library(simmer.plot)
make_parts <- trajectory("parts") %>% 
  set_attribute("start_time", function() {now(machineShop)}) %>% 
  seize("machine1") %>% 
  timeout(function() {rnorm(n = 1, mean = 10, sd = 1)}) %>% 
  release("machine1") %>% 
  seize("machine2", continue = FALSE, 
        reject = trajectory() %>%
          timeout(1) %>%
          rollback(amount = 2, times = Inf)) %>% 
  timeout(function() {rnorm(n = 1, mean = 5, sd = 2)}) %>% 
  release("machine2") %>% 
  seize("machine3", continue = FALSE, 
        reject = trajectory() %>%
          timeout(1) %>%
          rollback(amount = 2, times = Inf)) %>% 
  timeout(function() {runif(n = 1, min = 5, max = 15)}) %>% 
  release("machine3") %>% 
  seize("machine4", continue = FALSE, 
        reject = trajectory() %>%
          timeout(1) %>%
          rollback(amount = 2, times = Inf)) %>% 
  timeout(function() {runif(n = 1, min = 10, max = 15)}) %>% 
  release("machine4") %>% 
  seize("machine5", continue = FALSE, 
        reject = trajectory() %>%
          timeout(1) %>%
          rollback(amount = 2, times = Inf)) %>% 
  timeout(function() {rnorm(n = 1, mean = 10, sd = 2.5)}) %>% 
  release("machine5")

machineShop <- simmer("machineShop") %>% 
  add_resource("machine1", capacity = 1, queue_size = 1) %>%
  add_resource("machine2", capacity = 1, queue_size = 1) %>% 
  add_resource("machine3", capacity = 1, queue_size = 1) %>% 
  add_resource("machine4", capacity = 1, queue_size = 1) %>% 
  add_resource("machine5", capacity = 1, queue_size = 1) %>% 
  add_generator("part", make_parts, mon = 1, function() {c(0, rexp(1000, 1/.5), -1)})  

simmer::run(machineShop, 480)

result <- get_mon_arrivals(machineShop)

resourceResult <- get_mon_resources(machineShop)

plot(result, metric = "activity_time") +
  theme_minimal()

plot(result, metric = "waiting_time") +
  theme_minimal()

plot(resourceResult, metric="usage") +
  theme_minimal()

plot(resourceResult, metric="utilization") +
  theme_minimal()

sum(result$finished[result$finished == TRUE])
```

```{r}
make_parts <- trajectory("parts") %>% 
  set_attribute("start_time", function() {now(machineShop)}) %>% 
  seize("machine1") %>% 
  timeout(function() {rnorm(n = 1, mean = 10, sd = 1)}) %>% 
  release("machine1") %>% 
  seize("machine2", continue = FALSE, 
        reject = trajectory() %>%
          timeout(1) %>%
          rollback(amount = 2, times = Inf)) %>% 
  timeout(function() {rnorm(n = 1, mean = 5, sd = 2)}) %>% 
  release("machine2") %>% 
  seize("machine3", continue = FALSE, 
        reject = trajectory() %>%
          timeout(1) %>%
          rollback(amount = 2, times = Inf)) %>% 
  timeout(function() {runif(n = 1, min = 5, max = 15)}) %>% 
  release("machine3") %>% 
  seize("machine4", continue = FALSE, 
        reject = trajectory() %>%
          timeout(1) %>%
          rollback(amount = 2, times = Inf)) %>% 
  timeout(function() {runif(n = 1, min = 10, max = 15)}) %>% 
  release("machine4") %>% 
  seize("machine5", continue = FALSE, 
        reject = trajectory() %>%
          timeout(1) %>%
          rollback(amount = 2, times = Inf)) %>% 
  timeout(function() {rnorm(n = 1, mean = 10, sd = 2.5)}) %>% 
  release("machine5")

machineShop <- simmer("machineShop") %>% 
  add_resource("machine1", capacity = 1, queue_size = 1) %>%
  add_resource("machine2", capacity = 1, queue_size = 1) %>% 
  add_resource("machine3", capacity = 1, queue_size = 1) %>% 
  add_resource("machine4", capacity = 1, queue_size = 1) %>% 
  add_resource("machine5", capacity = 1, queue_size = 1) %>% 
  add_generator("part", make_parts, mon = 1, function() {c(0, rexp(1000, 1/.5), -1)})  

simmer::run(machineShop, 480)

result <- get_mon_arrivals(machineShop)
```